---
title: "Homework - I"
author: ' Sriram Kannan'
output:
  github_document
---


```{r}
## load prostate data
prostate <- 
  read.table(url(
    'https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))
```


```{r}
prostate
```


```{r}
## subset to training examples
prostate_train <- subset(prostate, train==TRUE)
```


```{r}
## plot lcavol vs lpsa
plot_psa_data <- function(dat=prostate_train) {
  plot(dat$lpsa, dat$lcavol,
       xlab="log Prostate Screening Antigen (psa)",
       ylab="log Cancer Volume (lcavol)",
       pch = 20)
}
plot_psa_data()
```


```{r}
## L2 loss function
L2_loss <- function(y, yhat)
  (y-yhat)^2
```

Write functions that implement the L1 loss and tilted absolute loss functions.'

```{r}

##L1 Loss function
L1_loss <- function(y, yhat)
  abs(y-yhat)

#Tilted Absolute Loss function

tilted_absolute_loss_t1 <- function(y, yhat, tau = 0.25)
  ifelse(y - yhat > 0, tau*(y-yhat), (tau-1)*(y-yhat))

tilted_absolute_loss_t2 <- function(y, yhat, tau = 0.75)
  ifelse(y - yhat > 0, tau*(y-yhat), (tau-1)*(y-yhat))

```


```{r}
## fit simple linear model using numerical optimization using L2 loss function
fit_lin <- function(y, x, loss=L2_loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
```

```{r}
## fit simple linear model using numerical optimization using L1 loss function
fit_lin_L1 <- function(y, x, loss=L1_loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}

## fit simple linear model using numerical optimization using Tilted Absolute loss function
fit_lin_TAL_t1 <- function(y, x, loss=tilted_absolute_loss_t1, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}

fit_lin_TAL_t2 <- function(y, x, loss=tilted_absolute_loss_t2, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
```


```{r}
## make predictions from linear model
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x
```


```{r}
## fit linear model (L2 Loss)
lin_beta <- fit_lin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L2_loss)

## fit linear model (L1 Loss)
lin_beta_L1 <- fit_lin_L1(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_loss)

## fit linear model (Tilted Absolute Loss)
lin_beta_TAL_t1 <- fit_lin_TAL_t1(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=tilted_absolute_loss_t1)

lin_beta_TAL_t2 <- fit_lin_TAL_t2(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=tilted_absolute_loss_t2)
```


```{r}
## compute predictions for a grid of inputs (L2_Loss)
x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_L2 <- predict_lin(x=x_grid, beta=lin_beta$par)

## compute predictions for a grid of inputs (L1_Loss)
lin_pred_L1 <- predict_lin(x=x_grid, beta=lin_beta_L1$par)

## compute predictions for a grid of inputs (Tilted_Absolute_Loss)
lin_pred_TAL_t1 <- predict_lin(x=x_grid, beta=lin_beta_TAL_t1$par)
lin_pred_TAL_t2 <- predict_lin(x=x_grid, beta=lin_beta_TAL_t2$par)

```

Create a figure that shows lpsa (x-axis) versus lcavol (y-axis). Add and label (using the 'legend' function) the linear model predictors associated with L2 loss, L1 loss, and tilted absolute value loss for tau = 0.25 and 0.75.

```{r}
plot_psa_data()
## plot predictions
lines(x=x_grid, y=lin_pred_L2, col='darkgreen', lwd=2)
lines(x=x_grid, y=lin_pred_L1, col='skyblue', lwd=2)
lines(x=x_grid, y=lin_pred_TAL_t1, col='pink', lwd=2)
lines(x=x_grid, y=lin_pred_TAL_t2, col='orange', lwd=2)
legend("topleft", legend = c("L2 Loss", "L1 Loss", "Tilted Abs. Loss (tau = 0.25)", "Tilted Abs. Loss (tau = 0.75)"), col = c("darkgreen", "skyblue", "pink", "orange"), lty = 1, lwd = 2)

```
Write functions to fit and predict from a simple nonlinear model with three parameters defined by 'beta[1] + beta[2]*exp(-beta[3]*x)'. Hint: make copies of 'fit_lin' and 'predict_lin' and modify them to fit the nonlinear model. Use c(-1.0, 0.0, -0.3) as 'beta_init'.

```{r}
predict_nonlin <- function(x, beta)
  beta[1] + beta[2]*exp(-beta[3]*x)
```

```{r}
## fit non linear model using numerical optimization using L2 loss function
fit_nonlin <- function(y, x, loss=L2_loss, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y, beta[1] + beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
```

```{r}
## fit non linear model using numerical optimization using L1 loss function
fit_nonlin_L1 <- function(y, x, loss=L1_loss, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}

## fit non linear model using numerical optimization using Tilted Absolute loss function
fit_nonlin_TAL_t1 <- function(y, x, loss=tilted_absolute_loss_t1, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}

fit_nonlin_TAL_t2 <- function(y, x, loss=tilted_absolute_loss_t2, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y, beta[1] + beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
```


```{r}
## fit non linear model (L2 Loss)
nonlin_beta <- fit_nonlin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L2_loss)

## fit non linear model (L1 Loss)
nonlin_beta_L1 <- fit_nonlin_L1(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_loss)

## fit non linear model (Tilted Absolute Loss)
nonlin_beta_TAL_t1 <- fit_nonlin_TAL_t1(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=tilted_absolute_loss_t1)

nonlin_beta_TAL_t2 <- fit_nonlin_TAL_t2(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=tilted_absolute_loss_t2)
```

```{r}
predict_nonlin <- function(x, beta)
  beta[1] + beta[2]*exp(-beta[3]*x)
```

```{r}
## compute predictions for a grid of inputs (L2_Loss)
x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
nonlin_pred_L2 <- predict_nonlin(x=x_grid, beta=nonlin_beta$par)

## compute predictions for a grid of inputs (L1_Loss)
nonlin_pred_L1 <- predict_nonlin(x=x_grid, beta=nonlin_beta_L1$par)

## compute predictions for a grid of inputs (Tilted_Absolute_Loss)
nonlin_pred_TAL_t1 <- predict_nonlin(x=x_grid, beta=nonlin_beta_TAL_t1$par)
nonlin_pred_TAL_t2 <- predict_nonlin(x=x_grid, beta=nonlin_beta_TAL_t2$par)

```

Create a figure that shows lpsa (x-axis) versus lcavol (y-axis). Add and label (using the 'legend' function) the nonlinear model predictors associated with L2 loss, L1 loss, and tilted absolute value loss for tau = 0.25 and 0.75.

```{r}
plot_psa_data()
## plot predictions
lines(x=x_grid, y=nonlin_pred_L2, col='darkgreen', lwd=2)
lines(x=x_grid, y=nonlin_pred_L1, col='skyblue', lwd=2)
lines(x=x_grid, y=nonlin_pred_TAL_t1, col='pink', lwd=2)
lines(x=x_grid, y=nonlin_pred_TAL_t2, col='orange', lwd=2)
legend("topleft", legend = c("L2 Loss", "L1 Loss", "Tilted Abs. Loss (tau = 0.25)", "Tilted Abs. Loss (tau = 0.75)"), col = c("darkgreen", "skyblue", "pink", "orange"), lty = 1, lwd = 2)

```