---
title: "Homework - 6"
author: ' Sriram Kannan'
output:
  github_document
---

```{r}
library(randomForest)
library(dplyr)
```


1) Convert the response variable in the “vowel.train” data frame to a factor variable prior to training, so that “randomForest” does classification rather than regression.

```{r}

```

2) Review the documentation for the “randomForest” function.

```{r}
?randomForest()
```

3) Fit the random forest model to the vowel data using all of the 11 features using the default values of the tuning parameters.

```{r}

```

4) Use 5-fold CV and tune the model by performing a grid search for the following tuning parameters: 1) the number of variables randomly sampled as candidates at each split; consider values 3, 4, and 5, and 2) the minimum size of terminal nodes; consider a sequence (1, 5, 10, 20, 40, and 80).

```{r}

```

5) With the tuned model, make predictions using the majority vote method, and compute the misclassification rate using the ‘vowel.test’ data.

```{r}

```

