---
title: "Homework - 6"
author: ' Sriram Kannan'
output:
  github_document
---

```{r}
library(randomForest)
library(dplyr)
library(magrittr)
library(caret)
```


Using the “vowel.train” data, and the “randomForest” function in the R package “randomForest”. Develop a random forest classifier for the vowel data by doing the following:

1) Convert the response variable in the “vowel.train” data frame to a factor variable prior to training, so that “randomForest” does classification rather than regression.

```{r}
vowel.train <- read.csv('voweltrain.csv')
vowel.train
vowel.train %<>%
  mutate_at(c('y'), ~as.factor(.)) 
```

2) Review the documentation for the “randomForest” function.

```{r}
?randomForest()
```

3) Fit the random forest model to the vowel data using all of the 11 features using the default values of the tuning parameters.

```{r}
fit <- randomForest(y ~ ., data=vowel.train)
```

4) Use 5-fold CV and tune the model by performing a grid search for the following tuning parameters: 1) the number of variables randomly sampled as candidates at each split; consider values 3, 4, and 5, and 2) the minimum size of terminal nodes; consider a sequence (1, 5, 10, 20, 40, and 80).

```{r}
#the number of variables randomly sampled as candidates at each split; consider values 3, 4, and 5, and 2 - mtry
#the minimum size of terminal nodes; consider a sequence (1, 5, 10, 20, 40, and 80) - nodesize
```

```{r}
set.seed(100)
mty_param = c(2,3,4,5)
nodesize_param = c(1, 5, 10, 20, 40, 80)
ct = trainControl(method = 'cv', number = 5, search = 'grid')
tunegrid = expand.grid(mtry = mty_param, min.node.size = nodesize_param, splitrule = "gini")
grid_fits = train(y ~ ., data = vowel.train, metric = 'Accuracy', method = 'ranger', tuneGrid = tunegrid, trControl = ct)
grid_fits
```


5) With the tuned model, make predictions using the majority vote method, and compute the misclassification rate using the ‘vowel.test’ data.

```{r}
vowel.test <- read.csv('voweltest.csv')
vowel.test %<>%
  mutate_at(c('y'), ~as.factor(.)) 

fit <- randomForest(y ~ ., data=vowel.train, mtry = 2, nodesize = 1, splitrule = "gini")
pred <- predict(fit, newdata = vowel.test)
confusionMatrix(pred, vowel.test$y)


```

