---
title: "Homework - 5"
author: ' Sriram Kannan'
output:
  github_document
---

1) Randomly split the mcycle data into training (75%) and validation (25%) subsets.

```{r}
library(MASS)
library(ISLR2)
set.seed(200)
index <- sample(133,0.75*133)
train <- mcycle[index, ]
test <- mcycle[-index, ]
train
test
```

2) Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values. (hint: the script already implements this)

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}

## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}

y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)
x_plot <- matrix(seq(min(x),max(x),length.out=133),133,1)
k_param = c(1,3,5,9,11,15,20,25)

for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
    lines(x_plot, yhat_plot, col="#882255", lwd=2)
  }

```

3) With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2
```


```{r}
## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

y <- train$accel
x <- matrix(train$times, length(train$times), 1)
x_plot = matrix(seq(min(x),max(x),length.out=99),99,1)
v = c()
#Each value is tuned to a specific value of K as seen above. 
for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    v[i] = error(y, yhat_plot)
}
v<-v[!is.na(v)]
plot(k_param, v, xlab="k", ylab="train error", type = "l", main = "Train error using Squared Error Loss")
```


```{r}
## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}



aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

df = c()
for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    df[i] = sum(diag(attr(yhat_plot, 'k')))
}
df <- df[!is.na(df)]
for (i in length(k_param))
  {
    v[i] = aic(y, yhat_plot, d = df[i])
}
plot(k_param, v, xlab="k", ylab="aic of training data", type = "l", main = "AIC Value of Training Data using Squared Error Loss")
```


```{r}
## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d

for (i in length(k_param))
  {
    v[i] = bic(y, yhat_plot, d = df[i])
}
plot(k_param, v, xlab="k", ylab="bic of training data", type = "l", main = "BIC Value of Training Data using Squared Error Loss")
```

```{r}
y <- test$accel
x <- matrix(test$times, length(test$times), 1)
x_plot = matrix(seq(min(x),max(x),length.out=34),34,1)
v = c()
#Each value is tuned to a specific value of K as seen above.
for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    v[i] = error(y, yhat_plot)
}
v<-v[!is.na(v)]
plot(k_param, v, xlab="k", ylab="test error", type = "l", main = "Test error using Squared Error Loss")
```

4) For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}
library(caret)
library(matrixStats)
set.seed(100)
flds <- createFolds(mcycle$accel, k = 5)
print(flds)
sapply(flds, length)

cvknn <- function(kNN = 10, flds = flds)
{
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds))
  {
    acc_trn <- mcycle[-flds[[tst_idx]],]
    acc_tst <- mcycle[flds[[tst_idx]],]
    
    knn_fit <- knnreg(accel ~ times, k = kNN, data = acc_trn)
    pre_tst <- predict(knn_fit, acc_tst)
    cverr[tst_idx] <- mean((acc_tst$accel - pre_tst)^2)
  }
  return(cverr)
}

cvvers = matrix(, nrow = length(k_param), ncol = 5)

for (i in 1:length(k_param))
{
  val <- cvknn(k_param[i], flds)
  cvvers[i,] <- val
}
cvvers
cvvers_mean <- rowMeans(cvvers)
as.matrix(cvvers_mean)
cvvers_sd <- rowSds(cvvers)
as.matrix(cvvers_sd)
```

5)Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}
plot(k_param, cvvers_mean, xlab="k", ylab="test error - 5-Fold Average", type = "l", main = "Average 5-Fold CV Error using Squared Error Loss")
sd <- 1
segments(k_param,cvvers_mean-sd,k_param,cvvers_mean+sd)
```

6) Interpret the resulting figures and select a suitable value for the tuning parameter.

The resulting figure shows a drastic reduction in average test error as we approach 5 for the tuning parameter with a slight improvement at k = 10 with the error increasing dramatically beyond 10. This result is expected as the model is overfit under k = 4 and is undergit beyond k = 13 or so. Hence, ideally we'd like k to be around 10.