---
title: "Homework - 5"
author: ' Sriram Kannan'
output:
  github_document
---

```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'

y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)

x
#plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
```


```{r}
## Epanechnikov kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## lambda - bandwidth (neighborhood size)
kernel_epanechnikov <- function(x, x0, lambda=1) {
  d <- function(t)
    ifelse(t <= 1, 3/4*(1-t^2), 0)
  z <- t(t(x) - x0)
  d(sqrt(rowSums(z*z))/lambda)
}
```


```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```


```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}
```


```{r}
## Helper function to view kernel (smoother) matrix
matrix_image <- function(x) {
  rot <- function(x) t(apply(x, 2, rev))
  cls <- rev(gray.colors(20, end=1))
  image(rot(x), col=cls, axes=FALSE)
  xlb <- pretty(1:ncol(x))
  xat <- (xlb-0.5)/ncol(x)
  ylb <- pretty(1:nrow(x))
  yat <- (ylb-0.5)/nrow(x)
  axis(3, at=xat, labels=xlb)
  axis(2, at=yat, labels=ylb)
  mtext('Rows', 2, 3)
  mtext('Columns', 3, 3)
}
```


```{r}
## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}
```


```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2
```


```{r}
## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))
```


```{r}
## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d
```


```{r}
## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d
```


```{r}


## how does k affect shape of predictor and eff. df using k-nn kernel ?
k_slider = kparam
manipulate({
  ## make predictions using NW method at training inputs
  y_hat <- nadaraya_watson(y, x, x,
    kern=kernel_k_nearest_neighbors, k=k_slider)
  edf <- effective_df(y, x,
    kern=kernel_k_nearest_neighbors, k=k_slider)
  aic_ <- aic(y, y_hat, edf)
  bic_ <- bic(y, y_hat, edf)
  y_hat_plot <- nadaraya_watson(y, x, x_plot,
    kern=kernel_k_nearest_neighbors, k=k_slider)
  plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
  legend('topright', legend = c(
    paste0('eff. df = ', round(edf,1)),
    paste0('aic = ', round(aic_, 1)),
    paste0('bic = ', round(bic_, 1))),
    bty='n')
  lines(x_plot, y_hat_plot, col="#882255", lwd=2)
}, k_slider=slider(1, 10, initial=3, step=1))

```

1) Randomly split the mcycle data into training (75%) and validation (25%) subsets.

```{r}
library(MASS)
library(ISLR2)
set.seed(200)
index <- sample(133,0.75*133)
train <- mcycle[index, ]
test <- mcycle[-index, ]
train
test
```

2) Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values. (hint: the script already implements this)

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}

## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}

y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)
x_plot <- matrix(seq(min(x),max(x),length.out=133),133,1)
k_param = c(1,3,5,9,11,15,20,25)

for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
    lines(x_plot, yhat_plot, col="#882255", lwd=2)
  }

```

3) With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2
```


```{r}
## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

y <- train$accel
x <- matrix(train$times, length(train$times), 1)
x_plot = matrix(seq(min(x),max(x),length.out=99),99,1)
v = c()
#Each value is tuned to a specific value of K as seen above. 
for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    v[i] = error(y, yhat_plot)
}
v<-v[!is.na(v)]
plot(k_param, v, xlab="k", ylab="train error", type = "l", main = "Train error using Squared Error Loss")
```


```{r}
## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}



aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

df = c()
for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    df[i] = sum(diag(attr(yhat_plot, 'k')))
}
df <- df[!is.na(df)]
for (i in length(k_param))
  {
    v[i] = aic(y, yhat_plot, d = df[i])
}
plot(k_param, v, xlab="k", ylab="aic of training data", type = "l", main = "AIC Value of Training Data using Squared Error Loss")
```


```{r}
## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d

for (i in length(k_param))
  {
    v[i] = bic(y, yhat_plot, d = df[i])
}
plot(k_param, v, xlab="k", ylab="bic of training data", type = "l", main = "BIC Value of Training Data using Squared Error Loss")
```

```{r}
y <- test$accel
x <- matrix(test$times, length(test$times), 1)
x_plot = matrix(seq(min(x),max(x),length.out=34),34,1)
v = c()
#Each value is tuned to a specific value of K as seen above.
for (i in k_param)
  {
    yhat_plot = nadaraya_watson(y = y,x = x,x0 = x_plot ,kern = kernel_k_nearest_neighbors, k = i)
    v[i] = error(y, yhat_plot)
}
v<-v[!is.na(v)]
plot(k_param, v, xlab="k", ylab="test error", type = "l", main = "Test error using Squared Error Loss")
```

4) For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}
library(caret)
library(glmnet)
?train()
```



5)Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}

```

6) Interpret the resulting figures and select a suitable value for the tuning parameter.