---
title: "Homework - 4"
author: ' Sriram Kannan'
output:
  github_document
---
When the number of features p is large, there tends to be a deterioration in the performance of KNN and other local approaches that perform prediction using only observations that are near the test observation for which a prediction must be made. This phenomenon is known as the curse of dimensionality, and it ties into the fact that parametric approaches often perform poorly when p is large. We will now investigate this curse.

(a) Suppose that we have a set of observations, each with measurements on p = 1 feature, X. We assume that X is uniformly (evenly) distributed on [0, 1]. Associated with each observation is a response value. Suppose that we wish to predict a test observation’s response using only observations that are within 10% of the range of X closest to that test observation. For instance, in order to predict the response for a test observation with X = 0.6, we will use observations in the range [0.55, 0.65]. On average, what fraction of the available observations will we use to make the prediction?

```{r}
(10*95 + 9 + 8 + 7 + 6 + 5)/100
```
On average we use 9.85% of the observations as over 0.95 or 95% of the given domain [0,1], we lose values on the right hand side of X due to 1 being a hard limit. 

(b) Now suppose that we have a set of observations, each with measurements on p = 2 features, X1 and X2. We assume that (X1,X2) are uniformly distributed on [0, 1] × [0, 1]. We wish to predict a test observation’s response using only observations that are within 10% of the range of X1 and within 10% of the range
of X2 closest to that test observation. For instance, in order to predict the response for a test observation with X1 = 0.6 and X2 = 0.35, we will use observations in the range [0.55, 0.65] for X1 and in the range [0.3, 0.4] for X2. On average, what fraction of the available observations will we use to make the prediction?

```{r}
sqrt(9.85)
```
We have doubled the dimensions while having the same number of points. Here we consider 2-D points with co-ordinates X1 and X2. Consider a hypercube or a hypersphere - the volume of the space is increases exponentially with increase in dimensions. We thus only use 3.13% of the observations when we have 2 dimensions. 

(c) Now suppose that we have a set of observations on p = 100 features. Again the observations are uniformly distributed on each feature, and again each feature ranges in value from 0 to 1. We wish to predict a test observation’s response using observations within the 10% of each feature’s range that is closest to that test observation. What fraction of the available observations will we use to make the prediction?

```{r}
9.85^(1/100)
```
Similarly, we only use 1.02% of the observations when we have 100 dimensions.

(d) Using your answers to parts (a)–(c), argue that a drawback of KNN when p is large is that there are very few training observations “near” any given test observation.

As can be seen in the previous answers, we're using less and less fractions of available observations precisely due to the fact that the density of observations exponentially decrease as the volume of the "hyperspace - i.e, space in n dimensions" increases. Thereby we conclude that there are very few training observations "near" any test observation. The "nearness" becomes harder and harder to discern as dimensions increase as explained further below.


(e) Now suppose that we wish to make a prediction for a test observation by creating a p-dimensional hypercube centered around the test observation that contains, on average, 10% of the training observations. For p = 1, 2, and 100, what is the length of each side of the hypercube? Comment on your answer.
Note: A hypercube is a generalization of a cube to an arbitrary number of dimensions. When p = 1, a hypercube is simply a line segment, when p = 2 it is a square, and when p = 100 it is a 100-dimensional cube.

Edge Length of Hypercube ep = f^1/p 
Plugging in the values, we have 
f = 0.1 (As test observations contain 10% of the training observations)

```{r}
e1 = 0.1^(1/1)
e2 = 0.1^(1/2)
e3 = 0.1^(1/100)
e1
e2
e3
```
As we can see, the length of the hypercube increases as the dimensions increase. Considering the maximum possible length is 1 (While using all observations in 1 dimension), we see that in order to utilize 10% of the training observations, in higher dimensions we'd have to search disproportionately farther away in order to find the values we need. There's a lot of empty space in high dimensions and hence it's very difficult to find observations that are close by. In case of 100 dimensions, we have to use 97.7% of the maximum length of the hypercube to get merely 10% of the observations showing how much farther apart the observations are. The key observation here is that the scaling is severly non-linear. In order to get 20% of the observations, we'd have to use 98.4% of the length and the difference decreases as we increase the % of observations used. The huge drawback here is that there's little difference between the closest point and the other points that are farther away. "Nearest Neighbours" loses its meaning to a large degree as the degree of "nearness" massively reduces as dimensions increase and becomes increasingly hard to ascertain when we reduce the fraction of neighbours we consider as "nearest". 


13. This question should be answered using the Weekly data set, which is part of the ISLR2 package. This data is similar in nature to the Smarket data from this chapter’s lab, except that it contains 1, 089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.

```{r}
library(ISLR2)
```


(a) Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?

```{r}
summary(Weekly)
```
```{r}
cor(Weekly[-9])
```
The only high correlation that is observed is between Year and Volume. The plot below shows this trend. The rise in index corresponds to the passage of weeks over time and hence is a time series starting from 1990 - 2009 in Week Numbers. 

```{r}
attach(Weekly)
plot(Volume)
```

(b) Use the full data set to perform a logistic regression with Direction as the response and the five lag variables plus Volume as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant? If so, which ones?

```{r}
glm.fits <- glm(
Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume ,
data = Weekly , family = binomial)
summary(glm.fits)
```
Based on the P-Values, We can conclude that Lag 2 is statistically significant as it has a p-value < 0.05. 

(c) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.

```{r}
glm.probs <- predict(glm.fits , type = "response")
glm.pred <- rep("Down", 1089)
glm.pred[glm.probs > .5] = "Up"
table(glm.pred , Direction)
mean(glm.pred == Direction)
```
The confusion matrix shows both the false positives - The number of predictions that were Down but were predicted as Up = 48 and the false negatives, i.e the number of Predictions that were Up but predicted as Down = 430.

(d) Now fit the logistic regression model using a training data period from 1990 to 2008, with Lag2 as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).

```{r}
train <- (Year <= 2008)
Weekly.2008 = Weekly[!train, ]
Direction.2008 <- Direction[!train]

glm.fits <- glm(
Direction ~ Lag2 ,
data = Weekly , family = binomial , subset = train
)
glm.probs <- predict(glm.fits , Weekly.2008,
type = "response")

glm.pred <- rep("Down", 104)
glm.pred[glm.probs > .5] <- "Up"
table(glm.pred , Direction.2008)

mean(glm.pred != Direction.2008)
```
The overall fraction of correct predictions for the held out data/test data is 0.375 in case of Logistic Regression. 

(e) Repeat (d) using LDA.

```{r}
library(MASS)
lda.fit <- lda(Direction ~ Lag2 , data = Weekly , subset = train)
lda.pred <- predict(lda.fit , Weekly.2008)
lda.class <- lda.pred$class
table(lda.class, Direction.2008)
mean(lda.class == Direction.2008)

```
The overall fraction of correct predictions for the held out data/test data is 0.625 in case of LDA. 

(f) Repeat (d) using QDA.

```{r}
qda.fit <- qda(Direction ~ Lag2 , data = Weekly , subset = train)
qda.pred <- predict(qda.fit , Weekly.2008)
qda.class <- qda.pred$class
table(qda.class, Direction.2008)
mean(qda.class == Direction.2008)
```
The overall fraction of correct predictions for the held out data/test data is 0.5865 in case of QDA. 

(g) Repeat (d) using KNN with K = 1.

```{r}
library(class)
library(tidyverse)
```

```{r}
train.X <- Weekly %>% filter(Year <= 2008) %>% select(Lag2)
test.X <- Weekly %>% filter(Year > 2008) %>% select(Lag2)
train.Direction <- Direction[train]
set.seed (1)
knn.pred <- knn(train.X, test.X, train.Direction , k = 1)
table(knn.pred , Direction.2008)
mean(knn.pred == Direction.2008)
```
The overall fraction of correct predictions for the held out data/test data is 0.5 in case of KNN using K = 1.

(h) Repeat (d) using naive Bayes.

```{r}
library(e1071)
nb.fit <- naiveBayes(Direction ~ Lag2 , data = Weekly, subset = train)
nb.class <- predict(nb.fit , Weekly.2008)
table(nb.class , Direction.2008)
mean(nb.class == Direction.2008)

```
The overall fraction of correct predictions for the held out data/test data is 0.5865 in case of Naive Bayes. 

(i) Which of these methods appears to provide the best results on this data?

Based on the Parameters used, for the Weekly Data using Lag2 as the sole predictor, LDA appears to provide the best results. 

(j) Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for
K in the KNN classifier.

```{r}
#Logistic Regression  - Using Lag1, 2 and 4 (Least 3 P-values)
glm.fits <- glm(
Direction ~  Lag1 + Lag2 - Lag4,
data = Weekly , family = binomial , subset = train
)
glm.probs <- predict(glm.fits , Weekly.2008,
type = "response")

glm.pred <- rep("Down", 104)
glm.pred[glm.probs > .5] <- "Up"
table(glm.pred , Direction.2008)

mean(glm.pred != Direction.2008)
```
We get a better prediction with Logistic Regression now, with using Lag 1, 2 and 4 in a different interaction that just using Lag 2. It still however, is lower than LDA with just Lag2. 

```{r}
#QDA
qda.fit <- qda(Direction ~ Lag1 + Lag2 - Lag4 , data = Weekly , subset = train)
qda.pred <- predict(qda.fit , Weekly.2008)
qda.class <- qda.pred$class
table(qda.class, Direction.2008)
mean(qda.class == Direction.2008)
```
We see a deterioration in the performance of QDA when using the same interaction with Lag1, Lag2 and Lag4 as above over the original model with just Lag2. 

```{r}
#Naive Bayes
nb.fit <- naiveBayes(Direction ~ Lag1 + Lag2 - Lag4 , data = Weekly, subset = train)
nb.class <- predict(nb.fit , Weekly.2008)
table(nb.class , Direction.2008)
mean(nb.class == Direction.2008)
```
We see a deterioration in the performance of Naive Bayes when using the same interaction with Lag1, Lag2 and Lag4 as above over the original model with just Lag2.


```{r}
train.X <- Weekly %>% filter(Year <= 2008) %>% select(Lag2)
test.X <- Weekly %>% filter(Year > 2008) %>% select(Lag2)
train.Direction <- Direction[train]
set.seed (1)
knn.pred <- knn(train.X, test.X, train.Direction , k = 20)
table(knn.pred , Direction.2008)
mean(knn.pred == Direction.2008)

```
In case of KNN, Using K = 20, We see an increase in model accuracy as opposed to using K = 1 with the same Lag2 as the predictor. Is is still however, lower than the accuracy of the original LDA model.

```{r}
lda.fit <- lda(Direction ~ Lag1 + Lag2 - Lag4 , data = Weekly , subset = train)
lda.pred <- predict(lda.fit , Weekly.2008)
lda.class <- lda.pred$class
table(lda.class, Direction.2008)
mean(lda.class == Direction.2008)
```
We see a deterioration in the performance of LDA when using the same interaction with Lag1, Lag2 and Lag4 as above over the original model with just Lag2.


We thus conclude that the original LDA model with an accuracy of 0.625 provides the best predictions with just Lag2 being used as the predictor. Which is understandable as that's the only statistically significant predictor based on the P-values with only Logistic Regression breaking this trend but still falling lower than LDA. 